Active Learning for SAT Solver Benchmarking {\large (extended and revised version)}

### 2024-05-25 paper_jar/paper.tex
Adapt names of prediction models
- remove hyphens
- spell out abbreviations

As a baseline for improvement, Table~\ref{tab:regression} shows the runtime prediction performance of two regression models, a random forest regressor and an MLP regressor, using 46 instance features and 27 known solver runtimes.
Both models were chosen for their robustness and ability to capture nonlinear dependencies~\cite{breiman2001random}, and both perform poorly, as evidenced by the magnitudes of the mean regression errors.
An RMSE error of about \SI{2000}{s} is quite poor compared to solver runtimes that range from 0 to \SI{5000}{s}.
Therefore, we decided to discretize the runtimes.
%
\begin{table}[b]
  \centering
  \caption{Regression performance of models trained on 46 instance and 27 runtime features. We report the RMSE and MAE metrics on the test sets.}
  \label{tab:regression}
  \begin{tabular}{
    >{\raggedleft\arraybackslash}m{0.33855\textwidth}
    >{\centering\arraybackslash}m{0.01\textwidth}
    >{\centering\arraybackslash}m{0.2562\textwidth}
    >{\centering\arraybackslash}m{0.2562\textwidth}
  }
    \toprule
    {Regression Models} & & {Avg. RMSE ($\pm$ std)} & {Avg. MAE ($\pm$ std)} \\
    \midrule
    Random-Forest-Regressor & & \SI{1929.84}{s} ($\pm$ \SI{1161.15}{s}) & \phantom{0}\SI{957.36}{s} ($\pm$ \SI{297.21}{s}) \\[0.4ex]
    MLP-Regressor           & & \SI{2275.20}{s} ($\pm$ \SI{1545.01}{s}) & \SI{1284.30}{s} ($\pm$ \SI{390.57}{s}) \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{table}[tbp]
  \centering
  \caption{MCC performance of models predicting timeouts. The models are trained 28 times with each solver as a target once.}
  \label{tab:timeout-prediction}
  ~\\[1em]
  \begin{tabular}{
    >{\raggedleft\arraybackslash}m{0.33855\textwidth}
    >{\centering\arraybackslash}m{0.01\textwidth}
    >{\centering\arraybackslash}m{0.2562\textwidth}
    >{\centering\arraybackslash}m{0.2562\textwidth}
  }
    \toprule
    {Timeout Prediction Models} & & {46 Instance Features Avg. MCC ($\pm$ std)} & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} \\
    \midrule
    Stacking (QDA\,+\,RF)                 & & 0.6513 ($\pm$ 0.0458) & 0.9527 ($\pm$ 0.0292) \\[0.4ex]
    Quadratic-DA (QDA)                    & & 0.2593 ($\pm$ 0.1310) & 0.9290 ($\pm$ 0.0339) \\[0.4ex]
    Random-Forest-Classifier (RF)         & & 0.6607 ($\pm$ 0.0547) & 0.8530 ($\pm$ 0.0479) \\[0.4ex]
    Ada-Boost                             & & 0.5412 ($\pm$ 0.0612) & 0.8384 ($\pm$ 0.0444) \\[0.4ex]
    Decision-Tree                         & & 0.5980 ($\pm$ 0.0423) & 0.8059 ($\pm$ 0.0707) \\[0.4ex]
    Logistic-Regression                   & & 0.2031 ($\pm$ 0.0728) & 0.8052 ($\pm$ 0.1018) \\[0.4ex]
    $k$NN-Classifier                      & & 0.5108 ($\pm$ 0.0387) & 0.7885 ($\pm$ 0.1521) \\[0.4ex]
    MLP-Classifier                        & & 0.1293 ($\pm$ 0.0718) & 0.7760 ($\pm$ 0.1408) \\[0.4ex]
    Support-Vector-Machines               & & 0.0595 ($\pm$ 0.0554) & 0.7757 ($\pm$ 0.2149) \\[0.4ex]
    Naive-Bayes                           & & 0.1173 ($\pm$ 0.0872) & 0.7306 ($\pm$ 0.1394) \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{table}[tbp]
  \centering
  \caption{MCC performance of the best models predicting the $C_k$ clusterings.}
  \label{tab:cluster-prediction}
  \begin{tabular}{
    >{\raggedleft\arraybackslash}m{0.33855\textwidth}
    >{\centering\arraybackslash}m{0.01\textwidth}
    >{\centering\arraybackslash}m{0.2562\textwidth}
    >{\centering\arraybackslash}m{0.2562\textwidth}
  }
    \toprule
    {$C_3$ Cluster Prediction Models} & & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
    \midrule
    Stacking (QDA\,+\,RF)                 & & 0.7464 ($\pm$ 0.0497) & 0.8380 ($\pm$ 0.0634) \\[0.4ex]
    Quadratic-DA (QDA)                    & & 0.6903 ($\pm$ 0.0607) & 0.7738 ($\pm$ 0.0459) \\[0.4ex]
    Random-Forest-Classifier (RF)         & & 0.7116 ($\pm$ 0.0385) & 0.7841 ($\pm$ 0.0469) \\[0.4ex]
    \midrule
    {$C_4$ Cluster Prediction Models} & & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
    \midrule
    Stacking (QDA\,+\,RF)                 & & 0.6562 ($\pm$ 0.0538) & 0.7366 ($\pm$ 0.0377) \\[0.4ex]
    Quadratic-DA (QDA)                    & & 0.6007 ($\pm$ 0.0765) & 0.6872 ($\pm$ 0.0469) \\[0.4ex]
    Random-Forest-Classifier (RF)         & & 0.6033 ($\pm$ 0.0455) & 0.6828 ($\pm$ 0.0527) \\[0.4ex]
    \midrule
    {$C_5$ Cluster Prediction Models} & & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
    \midrule
    Stacking (QDA\,+\,RF)                 & & 0.5835 ($\pm$ 0.0615) & 0.6396 ($\pm$ 0.0553) \\[0.4ex]
    Quadratic-DA (QDA)                    & & 0.5508 ($\pm$ 0.0669) & 0.5881 ($\pm$ 0.0912) \\[0.4ex]
    Random-Forest-Classifier (RF)         & & 0.5420 ($\pm$ 0.0459) & 0.5895 ($\pm$ 0.0475) \\
    \bottomrule
  \end{tabular}
\end{table}

>>>>

As a baseline for improvement, Table~\ref{tab:regression} shows the runtime prediction performance of two regression models, a random forest and a multilayer perceptron, using 46 instance features and 27 known solver runtimes.
Both models were chosen for their robustness and ability to capture nonlinear dependencies~\cite{breiman2001random}, and both perform poorly, as evidenced by the magnitudes of the mean regression errors.
An RMSE error of about \SI{2000}{s} is quite poor compared to solver runtimes that range from 0 to \SI{5000}{s}.
Therefore, we decided to discretize the runtimes.
%
\begin{table}[b]
  \centering
  \caption{Regression performance of models trained on 46 instance and 27 runtime features. We report the RMSE and MAE metrics on the test sets.}
  \label{tab:regression}
  \begin{tabular}{
    >{\raggedleft\arraybackslash}m{0.33855\textwidth}
    >{\centering\arraybackslash}m{0.01\textwidth}
    >{\centering\arraybackslash}m{0.2562\textwidth}
    >{\centering\arraybackslash}m{0.2562\textwidth}
  }
    \toprule
    {Regression Models} & & {Avg. RMSE ($\pm$ std)} & {Avg. MAE ($\pm$ std)} \\
    \midrule
    Random Forest & & \SI{1929.84}{s} ($\pm$ \SI{1161.15}{s}) & \phantom{0}\SI{957.36}{s} ($\pm$ \SI{297.21}{s}) \\[0.4ex]
    Multilayer Perceptron & & \SI{2275.20}{s} ($\pm$ \SI{1545.01}{s}) & \SI{1284.30}{s} ($\pm$ \SI{390.57}{s}) \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{table}[tbp]
  \centering
  \caption{MCC performance of models predicting timeouts. The models are trained 28 times with each solver as a target once.}
  \label{tab:timeout-prediction}
  ~\\[1em]
  \begin{tabular}{
    >{\raggedleft\arraybackslash}m{0.4\textwidth}
    >{\centering\arraybackslash}m{0.25\textwidth}
    >{\centering\arraybackslash}m{0.25\textwidth}
  }
    \toprule
    {Timeout Prediction Models} & {46 Instance Features Avg. MCC ($\pm$ std)} & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} \\
    \midrule
    Stacking (QDA\,+\,RF)                 & 0.6513 ($\pm$ 0.0458) & 0.9527 ($\pm$ 0.0292) \\[0.4ex]
    Quadratic Discriminant Analysis (QDA)                    & 0.2593 ($\pm$ 0.1310) & 0.9290 ($\pm$ 0.0339) \\[0.4ex]
    Random Forest (RF)         & 0.6607 ($\pm$ 0.0547) & 0.8530 ($\pm$ 0.0479) \\[0.4ex]
    AdaBoost                             & 0.5412 ($\pm$ 0.0612) & 0.8384 ($\pm$ 0.0444) \\[0.4ex]
    Decision Tree                         & 0.5980 ($\pm$ 0.0423) & 0.8059 ($\pm$ 0.0707) \\[0.4ex]
    Logistic Regression                   & 0.2031 ($\pm$ 0.0728) & 0.8052 ($\pm$ 0.1018) \\[0.4ex]
    $k$NN                   & 0.5108 ($\pm$ 0.0387) & 0.7885 ($\pm$ 0.1521) \\[0.4ex]
    Multilayer Perceptron                      & 0.1293 ($\pm$ 0.0718) & 0.7760 ($\pm$ 0.1408) \\[0.4ex]
    Support Vector Machine               & 0.0595 ($\pm$ 0.0554) & 0.7757 ($\pm$ 0.2149) \\[0.4ex]
    Naive Bayes                           & 0.1173 ($\pm$ 0.0872) & 0.7306 ($\pm$ 0.1394) \\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{table}[tbp]
  \centering
  \caption{MCC performance of the best models predicting the $C_k$ clusterings.}
  \label{tab:cluster-prediction}
  \begin{tabular}{
    >{\raggedleft\arraybackslash}m{0.4\textwidth}
    >{\centering\arraybackslash}m{0.25\textwidth}
    >{\centering\arraybackslash}m{0.25\textwidth}
  }
    \toprule
    {$C_3$ Cluster Prediction Models} & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
    \midrule
    Stacking (QDA\,+\,RF)                 & 0.7464 ($\pm$ 0.0497) & 0.8380 ($\pm$ 0.0634) \\[0.4ex]
    Quadratic Discriminant Analysis (QDA)                     & 0.6903 ($\pm$ 0.0607) & 0.7738 ($\pm$ 0.0459) \\[0.4ex]
    Random Forest (RF)         & 0.7116 ($\pm$ 0.0385) & 0.7841 ($\pm$ 0.0469) \\[0.4ex]
    \midrule
    {$C_4$ Cluster Prediction Models} & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
    \midrule
    Stacking (QDA\,+\,RF)                 & 0.6562 ($\pm$ 0.0538) & 0.7366 ($\pm$ 0.0377) \\[0.4ex]
    Quadratic Discriminant Analysis (QDA)                     & 0.6007 ($\pm$ 0.0765) & 0.6872 ($\pm$ 0.0469) \\[0.4ex]
    Random Forest (RF)         & 0.6033 ($\pm$ 0.0455) & 0.6828 ($\pm$ 0.0527) \\[0.4ex]
    \midrule
    {$C_5$ Cluster Prediction Models} & {46\,+\,27 Runtime Feat. Avg. MCC ($\pm$ std)} & {46\,+\,27\,+\,27 Clus. Feat. Avg. MCC ($\pm$ std)} \\
    \midrule
    Stacking (QDA\,+\,RF)                 & 0.5835 ($\pm$ 0.0615) & 0.6396 ($\pm$ 0.0553) \\[0.4ex]
    Quadratic Discriminant Analysis (QDA)                      & 0.5508 ($\pm$ 0.0669) & 0.5881 ($\pm$ 0.0912) \\[0.4ex]
    Random Forest (RF)         & 0.5420 ($\pm$ 0.0459) & 0.5895 ($\pm$ 0.0475) \\
    \bottomrule
  \end{tabular}
\end{table}
